---
title: "Econ144 hw1_Sijia"
author: "Sijia Hua"
date: "4/8/2019"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
library(dplyr)
library(gdata)
library(moments)
library("ggpubr")
library(dynlm)
library(tidyr)
```

## Econ 144 HW1

### Chapter 2 #2
```{r Chapter 2 #2}
setwd("/Users/Renaissance/Desktop")
hw11<-read.xls('econ144hw1data.xlsx',sheet = 1) #import data
```

```{r}
# compute descriptive statistics 
 hw11 %>% select(c(-1)) %>% summary()
paste("The skewness of GRGDP is ", round(skewness(hw11$GRGDP),3),". Hence GRGDP is left-skewness.") 
paste("The skewness of Return is ", round(skewness(hw11$RETURN),3),". Hence Return is left-skewness.") 
paste("The kurtosis of GRGDP is ",round(kurtosis(hw11$GRGDP),3))
paste("The kurtosis of Return is ",round(kurtosis(hw11$RETURN),3))
paste("The Jarque-Bera Test of GRGDP is ")
jarque.test(hw11$GRGDP)
paste("The Jarque-Bera Test of Return is ")
jarque.test(hw11$RETURN)


hist(hw11$GRGDP,main="Histogram of GRGDP",xlab="GRGDP")
hist(hw11$RETURN, main="Histogram of SP500 Quarterly Return",xlab="Return")

```
```{r boxplot in 2}
boxplot(hw11$GRGDP,main="Box Plot of GRGDP",horizontal = TRUE)
boxplot(hw11$RETURN,main="Box Plot of Return",horizontal = TRUE)

```



```{r correlation in #2}
# ls <- lm(hw12$GRGDP~hw12$RETURN) #find linear regression between GDP and returns
# summary(ls)
paste("Correlation betwen this two variable is ",round(cor(hw11$RETURN,hw11$GRGDP,method="pearson"),3), ". This indicates that GRGDP has a low correlation with Return.")
```

### Chapter 2 #4
```{r Chapter 2 #4}
# still use hw11 data
y=ts(hw11$GRGDP,start=1950.25, freq=1)
x=ts(hw11$RETURN,start=1950.25, freq=1)

# (a) 
ma=lm(y~x)
summary(ma)

```


\indent I choose the significant level = 0.05. As the p-value is 1.6e-05, much less than 0.05, we reject the null hypothesis that β = 0. Hence, there is a (significant) relationship between the variables in the linear regression model of the data set. In this case, stock market is a leading indicator.  


```{r 2.4 model b}
# (b) 
mb=dynlm(y~L(x,1))
summary(mb)    #r-square has improved 
```

\indent I choose the significant level = 0.05. As the p-value is 3.17e-11, much less than 0.05, we reject the null hypothesis that β = 0. Hence, there is a (significant) relationship between the variables in the linear regression model of the data set. In this case, stock market is a leading indicator. 

### Chapter 2 #5
```{r model c in #5}
# (c) 
mc=dynlm(y~L(x,1)+L(x,2)+L(x,3)+L(x,4))
summary(mc)
```

\indent  In summary, The p value of L(x,1) is 1.95e-08, which is smaller than 0.05. We reject H0: \beta_1 =0. The stock market with 1-quarter leading is an indicator. The p value of L(x,2) is 0.0867, which is bigger than 0.05. We can not reject H0. There is no significant difference. Stock market's 2-quarter leading price may not be an indicator. The p value of L(x,3) and L(x,4) is 0.1366 and 0.2201 respectively, which are bigger than 0.05. We can not reject H0.There is no significant difference. <br>
\indent P value of the F-statistic is 2.509e-11, which is smaller than 0.05.Hence, we can reject H0, the overall addition of the variables is significantly improving the model. 

```{r test model d}
md=dynlm(y~L(x,1)+L(x,2)+L(x,3)+L(x,4)+L(y,1))
summary(md)
```


### Chapter 2 #7

```{r}
hw12<-read.xls('econ144hw1data.xlsx',sheet = 2) # read data

# calculate growth rate
grate_unem <- function(x)(x/lag(x)-1)*100
hw12$growth_unem <-round(grate_unem(hw12$UNEM),3)
grate_pov <- function(x)(x/lag(x)-1)*100
hw12$growth_pov <- round(grate_pov(hw12$POV),3)

# delete blank column
hw12<-hw12%>% select(c(-4))

# drop first row (NA)
newhw12 <- hw12 %>% drop_na()
```

### descriptive statistics of two growth rates
```{r}
# descriptive statistic
summary(hw12)
paste("The skewness of growth rate of unemployed persons is ", round(skewness(newhw12$growth_unem),3),". Hence the growth rate of unemployed person is right-skewness.") 
paste("The skewness of growth rate of number of people in poverty is ", round(skewness(newhw12$growth_pov),3),". Hence Return is right-skewness.") 
paste("The kurtosis of growth rate of unemployed person is ",round(kurtosis(newhw12$growth_unem),3))
paste("The kurtosis of growth rate of number of people in poverty is ",round(kurtosis(newhw12$growth_pov),3))
paste("The Jarque-Bera Test of growth rate of unemployed person is ")
jarque.test(newhw12$growth_unem)
paste("The Jarque-Bera Test of growth rate of number of people is ")
jarque.test(newhw12$growth_pov)

```

```{r histogram}
hist(hw12$growth_unem,main="Histogram of growth rate of unemployed perons",xlab="Growth rate of unemployment")
hist(hw12$growth_pov, main="Histogram of growth rate of number of people in poverty",xlab="Growth rate of people in poverty")
```

```{r boxplot}
boxplot(hw12$growth_unem,main="Box Plot of growth rate of unemployment",horizontal = TRUE)
boxplot(hw12$growth_pov,main="Box Plot of growth rate of poverty",horizontal = TRUE)
```

```{r correlation}
plot(hw12$year,hw12$growth_unem,type="l",lwd = 1.5,xlab = "year",ylab="Growth rate")
lines(hw12$year,hw12$growth_pov,type="l",lwd=1.5, col="blue")
legend("topleft", legend = c("g.r.of unemployment", "g.r.of poverty"), col = c("black","blue"), pch = c(1,1), bty = "n", pt.cex = 1, cex = 1, text.col = "black", horiz = F)

paste("Correlation betwen this two variable is ",round(cor(hw11$RETURN,hw11$GRGDP,method="pearson"),3), ". This indicates that GRGDP has a low correlation with Return.")

```

### Chapter 2 #8

```{r}
hw13<-read.xls('econ144hw1data.xlsx',sheet = 3) # read data
y=ts(hw13$G_POV,start=1960, freq=1)
x=ts(hw13$G_UNEM,start=1960, freq=1)

```

```{r model a}
# (a) 
modela=lm(y~x)
summary(modela)
```
For model a, multiple R-squared is 0.5045, adjusted R-squared is 0.4944.

```{r model b}
# I choose model b,c,d in excercise #3

# (b) 
modelb=dynlm(y~L(x,1))
summary(modelb)
```
\indent For model b, multiple R-squared is 0.1267, adjusted r-squared is 0.1086.
```{r model c}
# (c) 
modelc=dynlm(y~L(x,1)+L(x,2)+L(x,3)+L(x,4))
summary(modelc)
```
\indent For model c, multiple R-squared is 0.1596, adjusted R-squared is 0.07957.
```{r model d}
# (d) 
modeld=dynlm(y~L(x,1)+L(x,2)+L(x,3)+L(x,4)+L(y,1))
summary(modeld)
```
\indent For model d, multiple R-squared is 0.3694, adjusted R-squared is 0.2925. 

\indent <br> By comparing adjusted R-squared, I prefer model a. <br>
\indent Compare AIC and BIC:

```{r AIC,BIC}
AIC(modela,modelb,modelc,modeld)
BIC(modela,modelb,modelc,modeld)

```


\indent If I select model with respect to lowest AIC/BIC, I will prefer model a. 




